---
title: "Top Album Ranking Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("ggplot2")
library("reshape2")

aotd <- read_csv("./data/AOTD.csv", col_types = cols(
  X1 = col_skip(),
  artist = col_character(),
  album = col_character(),
  rank = col_integer(),
  genre = col_character(),
  year = col_integer(),
  reviewer_url = col_skip()
))


```

# Top Albums of the Decade 2010-2019

Back in December 2019 many music reporters released lists of their favorite albums of the decade.
I noticed some similarities and differences between different lists, and this led me to the
question I would like to explore in this report, which is 

<p style="text-align: center;"> How different are top decade album lists? </p>

# Methodology

I scraped the first few pages of Google for the top albums of the decade using the
Python package `Beautiful Soup`. The result is a dataframe that has the features 
`artist`, `album`, `genre`, `rank`, and `reviewer`. The data was scraped from 
Paste, Pitchfork, Stereogum, Consequence of Sound, Genius, Billboard, Rolling Stone,
and Time magazine.

Most of these reviewers ranked 100 albums. The exceptions are Pitchfork, who ranked 200 albums, and
Time, who ranked 10 albums without assigning a numeric value. I will treat these 10 albums
as all ranked equally in 10th place.



# Exploration

To deal with the problem of not every album appearing on every ranking, I borrowed
the scoring idea from football rankings, where personal rankings are replaced
by scores. So, if a reviewer gives player A the 1st place ranking, player B 2nd place ranking, and so on, then person A gets, say, 10 points added to their score, person B gets 9 points, and so on. 
I took a simple approach of subtracting the rank from the maximum possible rank (plus one), which was 200. That way first place gets 200 points, second gets 199, and so on. 


```{r pca}

# create scores
aotd$score <- 201 - aotd$rank

# create bins for the scores
aotd$top_group <- cut(aotd$rank, c(1, 10, 20, 30, 40, 50, 200), c(paste("Top", c(10, 20, 30, 40, 50)), "> 50"))

# long form to wide form
aotd_wide <- dcast(aotd, album + artist ~ reviewer, value.var="score", fill=0)
names(aotd_wide) <- gsub("_", " ", names(aotd_wide))

# pca
rev_pca <- prcomp(aotd_wide[-(1:2)])
autoplot(rev_pca, loadings=TRUE, loadings.label=TRUE)
```










