---
title: "Analysis of Top Albums of the Decade Rankings"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
library("tidyverse")
library("kableExtra")
library("ggplot2")
library("reshape2")
# library("ggfortify")
library("plotly")
library("ggbiplot")

aotd <- read_csv("./data/AOTD.csv", col_types = cols(
  X1 = col_skip(),
  artist = col_character(),
  album = col_character(),
  rank = col_integer(),
  genre = col_character(),
  year = col_integer(),
  reviewer_url = col_skip()
))
```

```{r preprocessing}

# create scores
aotd$score <- 201 - aotd$rank

# create bins for the scores
aotd$top_group <- cut(aotd$rank, c(1, 10, 20, 30, 40, 50, 200), c(paste("Top", c(10, 20, 30, 40, 50)), "> 50"))

# Long form to wide form
aotd_wide <- dcast(aotd, album + artist ~ reviewer, value.var="score", fill=0)
names(aotd_wide) <- gsub("_", " ", names(aotd_wide))

# number of lists each album appears on
aotd_wide$n_lists <- rowSums(aotd_wide[-(1:2)]!=0)


generate_PCA_plot <- function(album_df, title) {
  df_ranks <- album_df
  df_ranks[-(1:2)] <-
    (
      df_ranks[-(1:2)]
      %>% lapply(FUN= function(x) ifelse(x!=0, 201-x, NA))
    )
  
  # Generates the hovertext for the PCA plot
  hovertext <- apply(df_ranks,
                     MARGIN = 1,
                     FUN = function(row) {
                       paste(
                         paste0("</br>", 
                                colnames(df_ranks)
                                ), 
                         row, 
                         sep=": ", 
                         collapse = " ") 
                       }
                     )
  
  # PCA
  rev_pca <- (
    album_df
    %>% select(-album, -artist)
    %>% prcomp(scale=TRUE)
  )
  
  pre_plot <- ggbiplot(rev_pca, scale=0) +
    geom_point(aes(text=hovertext)) +
    ggtitle(title)
  
  plotly_plot <- ggplotly(pre_plot, tooltip="text") 
  
  return(plotly_plot)
}

```

Back in December 2019 many music reporters released lists of their favorite albums of the decade.
I noticed some similarities and differences between different lists, and this led me to two 
questions I would like to explore in this report:

1. How different are top decade album lists?
2. Is there consensus on the top albums of the decade?

# Methodology

I scraped the first few pages of Google for the top albums of the decade using the
Python package `Beautiful Soup`. The result is a dataframe that has the features 
`artist`, `album`, `genre`, `rank`, and `reviewer`. The data was scraped from 
Paste, Pitchfork, Stereogum, Consequence of Sound, Genius, Billboard, Rolling Stone,
and Time magazine. Due to each website having its own HTML formatting I had to do
the scraping nearly by hand.

Most of these reviewers ranked 100 albums. The exceptions are Pitchfork, who ranked 200 albums, and
Time, who ranked 10 albums without assigning a numeric value. I will treat these 10 albums
as all ranked equally in 10th place. In all there are 810 rows and 8 features in this data set.

```{r show_data}

kable(head(select(aotd, -top_group))) %>%
  kable_styling(bootstrap_options=c("striped", "hover", "condensed"))

```

To deal with the problem of not every album appearing on every ranking, I borrowed
a scoring idea from football rankings. The idea is to give a score to each rank. 
So, say a reviewer gives player A the 1st place ranking, player B 2nd place ranking, and so on, 
then person A gets 10 points added to their score, person B gets 9 points, and so on. 
I took a simple approach of subtracting the rank from the maximum possible rank (plus one), which was 200. That way first place gets 200 points, second gets 199, and so on. 

# Exploration

The first idea I had was to do Principal Component Analysis (PCA) on the data
to get a feel for the reviewers and the albums. This turned out to give 
a surprising amount of information.


## PCA on scores {.tabset}

To gain some insight on how albums are spaced, I cast the dataframe into wide format along the score feature.
This gave me a dataframe with albums in the rows and reviewers in the columns:

```{r show_wide_form}
aotd_wide %>% select(-n_lists) %>% head %>% kable %>% kable_styling(bootstrap_options=c("striped", "hover", "condensed"))
```

Here I would have a dataframe full of `NA`'s on albums that didn't get a ranking if I had not converted to scores. Now, any zero in this array means an album didn't appear on that website's list.

I performed a few variations on the PCA analysis, which you can tab through below. 
First I tried it with all of the albums, but you see many collinear data points which come from
low-ranked albums that only appear on one list. 
The second plot removes those albums and only shows albums appearing on more than one list.
The third plot shows the result of removing Time from the ranking, since only
10 albums were reviewed. 

Mouse over the plots to see more information about data points.

### With All Albums

```{r all_PCA_plot, fig.width=10, fig.height=8, warning=FALSE}

 aotd_wide %>% select(-n_lists) %>% generate_PCA_plot("PCA using all albums")

```

### Albums on Multiple Lists

```{r multiple_PCA_plot, fig.width=10, fig.height=8, warning=FALSE}

aotd_wide %>%
  dplyr::filter(n_lists>1) %>%
  select(-n_lists) %>%
  generate_PCA_plot("PCA using albums appearing only on more than one list")

```

### Multiple Lists, no Time

```{r PCA_no_Time, fig.width=10, fig.height=8, warning=FALSE}

aotd_wide %>%
  dplyr::filter(n_lists>1) %>%
  select(-n_lists, -Time) %>%
  generate_PCA_plot("PCA using albums appearing only on more than one list, excluding Time Magazine")

```

## PCA Interpretation

There seems to be a clear interpretation of the first two principal components in this setting:
the first component seems to be measuring average score, with higher-scoring albums off to the left
and lower-scoring albums off to the right.
The second component seems to be measuring consensus between the reviewers:
albums closer to the $x$-axis (i.e. with PC2 $\approx 0$) have a higher consensus between rankers,
and albums farther away from the x-axis seem to have lower consensus.

Based on the trends above there seems to be a few noteworthy clusters of albums. 

1. A cluster near the $x$-axis farthest to the left. Inspecting these albums tells me they 
all have high rankings.

2. Albums near the $y$-axis but far away from the x-axis. Inspecting tells me they
have mixed rankings.

3. Albums near the $x$-axis farthest to the right. These are albums that get
low scores from most reviewers.

It seems there is an answer here to the question "which albums are truly albums of the decade?"
This seems to be those in the first cluster.

# Future Directions

The loadings in the PCA plots above give some indication for how similar the ranking lists
are - the closer the arrows, the more similar the lists. I am curious about
finding a more clear method of deciding the similarity between rankers.
Furthermore, I have not yet used the genre feature in analyzing this data, and it would be interesting to see if
there are any interesting patterns that emerge if genre is included.





